COFFEE: an Optimizing Compiler\\for Finite Element Local Assembly

### 2014-04-14 paper/bare_conf.tex
Update plots and table

\begin{algorithm}[t]
\small
\KwSty{void} helmholtz(double A[3][3], double **coords) $\lbrace$\\
~~// K, det = Compute Jacobian (coords) \\
~~\\
~~\KwSty{static const double} W3[3] = $\lbrace$...$\rbrace$\\
~~\KwSty{static const double} X$\_$D10[3][3] = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\KwSty{static const double} X$\_$D01[3][3] = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$3; i++) \\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$3; j++) \\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$3; k++) \\
~~~~~~~~A[j][k] += ((Y[i][k]*Y[i][j]+\\
~~~~~~~~~~~+((K1*X$\_$D10[i][k]+K3*X$\_$D01[i][k])*\\
~~~~~~~~~~~*(K1*X$\_$D10[i][j]+K3*X$\_$D01[i][j]))+\\
~~~~~~~~~~~+((K0*X$\_$D10[i][k]+K2*X$\_$D01[i][k])*\\
~~~~~~~~~~~*(K0*X$\_$D10[i][j]+K2*X$\_$D01[i][j])))*\\
~~~~~~~~~~~*det*W3[i]);\\
$\rbrace$
\caption{Local assembly code generated by Firedrake for a Helmholtz problem on a 2D triangular mesh using Lagrange $p=1$ elements.}
\label{code:helmholtz}
\end{algorithm}

\begin{algorithm}[t]
\small
\KwSty{void} burgers(double A[12][12], double **c, double **w) $\lbrace$\\
~~// K, det = Compute Jacobian (c) \\
~~\\
~~\KwSty{static const double} W5[5] = $\lbrace$...$\rbrace$\\
~~\KwSty{static const double} X1$\_$D001[5][12] = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\KwSty{static const double} X2$\_$D001[5][12] = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~//11 other basis functions definitions.\\
~~...\\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$5; i++) $\lbrace$\\
~~~~\KwSty{double} F0 = 0.0;\\
~~~~//10 other declarations (F1, F2,...)\\
~~~~...\\
~~~~\KwSty{for} (\KwSty{int} r = 0; r$<$12; r++) $\lbrace$\\
~~~~~~F0 += (w[r][0]*X1$\_$D100[i][r]);\\
~~~~~~//10 analogous statements (F1, F2, ...)\\
~~~~$\rbrace$\\
~~~~...\\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$12; j++) \\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$12; k++) \\
~~~~~~~~A[j][k] += (..(K5*F9)+(K8*F10))*Y1[i][j])+\\
~~~~~~~~~+(((K0*X1$\_$D100[i][k])+(K3*X1$\_$D010[i][k])+\\
~~~~~~~~~+(K6*X1$\_$D001[i][k]))*Y2[i][j]))*F11)+\\
~~~~~~~~~+(..((K2*X2$\_$D100[i][k])+...+(K8*X2$\_$D001[i][k])*\\
~~~~~~~~~*(K2*X2$\_$D100[i][j])+...+(K8*X2$\_$D001[i][j]))..)+\\
~~~~~~~~~+ $<$roughly a hundred sum/muls go here$>$)..)*\\
~~~~~~~~~*det*W5[i]);\\
~~$\rbrace$ \\
$\rbrace$
\caption{Local assembly code generated by Firedrake for a Burgers problem on a 3D tetrahedral mesh using Lagrange $p=1$ elements.}
\label{code:burgers}
\end{algorithm}

@@

\begin{algorithm}[t]
\small
\KwSty{void} helmholtz(double A[3][4], double **coords) $\lbrace$\\
~~\KwSty{$\#$define} ALIGN $\_\_$attribute$\_\_$((aligned(32))) \\
~~// K, det = Compute Jacobian (coords) \\
~~\\
~~\KwSty{static const double} W3[3] ALIGN = $\lbrace$...$\rbrace$\\
~~\KwSty{static const double} X$\_$D10[3][4] ALIGN = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\KwSty{static const double} X$\_$D01[3][4] ALIGN = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$3; i++) $\lbrace$ \\
~~~~double LI$\_$0[4] ALIGN;\\
~~~~double LI$\_$1[4] ALIGN;\\
~~~~\KwSty{for} (\KwSty{int} r = 0; r$<$4; r++) $\lbrace$ \\
~~~~~~LI$\_$0[r] = ((K1*X$\_$D10[i][r])+(K3*X$\_$D01[i][r]));\\
~~~~~~LI$\_$1[r] = ((K0*X$\_$D10[i][r])+(K2*X$\_$D01[i][r]));\\
~~~~$\rbrace$\\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$3; j++) \\
~~~~~~\KwSty{$\#$pragma vector aligned}\\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k++) \\
~~~~~~~~A[j][k] += (Y[i][k]*Y[i][j]+LI$\_$0[k]*LI$\_$0[j]+\\
~~~~~~~~~~~~~~~~~~~~+LI$\_$1[k]*LI$\_$1[j])*det*W3[i]);\\
~~$\rbrace$\\
$\rbrace$
\caption{Local assembly code generated by Firedrake when padding, data alignment, and \emph{licm} are applied to the Helmholtz problem given in Figure~\ref{code:helmholtz} for an AVX architecture. In this specific case, sub-expressions invariant to $j$ are identical to those invariant to $k$, so they can be precomputed once in the $r$ loop.}
\label{code:helmholtz-licm}
\end{algorithm}

@@

\begin{algorithm}[t]
\small
\KwSty{void} helmholtz(double A[8][8], double **coords) $\lbrace$\\
~~// K, det = Compute Jacobian (coords) \\
~~// Declaration of basis function matrices \\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$6; i++) $\lbrace$ \\
~~~~// Do loop-invariant code motion \\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$4; j+=4) \\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$8; k+=4) $\lbrace$\\
~~~~~~~~// $load$ and $set$ intrinsics \\
~~~~~~~~// Compute A[1,1],A[2,2],A[3,3],A[4,4] \\
~~~~~~~~// One $permute\_pd$ intrinsics per $k$-loop $load$\\
~~~~~~~~// Compute A[1,2],A[2,1],A[3,4],A[4,3] \\
~~~~~~~~// One $permute2f128\_pd$ intrinsics per $k$-loop $load$\\
~~~~~~~~// ...\\
~~~~~~$\rbrace$\\
~~~~// Remainder loop (from $j=4$ to $j=6$)\\
~~$\rbrace$\\
~~// Restore the storage layout:\\
~~\KwSty{for} (\KwSty{int} j = 0; j$<$4; j+=4) $\lbrace$\\
~~~~$\_\_$m256d r0, r1, r2, r3, r4, r5, r6, r7;\\
~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$8; k+=4) $\lbrace$\\
~~~~~~r0 = $\_$mm256$\_$load$\_$pd ($\&$A[j+0][k]);\\
~~~~~~// Load A[j+1][k], A[j+2][k], A[j+3][k]\\
~~~~~~r4 = $\_$mm256$\_$unpackhi$\_$pd (r1, r0);\\
~~~~~~r5 = $\_$mm256$\_$unpacklo$\_$pd (r0, r1);\\
~~~~~~r6 = $\_$mm256$\_$unpackhi$\_$pd (r2, r3);\\
~~~~~~r7 = $\_$mm256$\_$unpacklo$\_$pd (r3, r2);\\
~~~~~~r0 = $\_$mm256$\_$permute2f128$\_$pd (r5, r7, 32);\\
~~~~~~r1 = $\_$mm256$\_$permute2f128$\_$pd (r4, r6, 32);\\
~~~~~~r2 = $\_$mm256$\_$permute2f128$\_$pd (r7, r5, 49);\\
~~~~~~r3 = $\_$mm256$\_$permute2f128$\_$pd (r6, r4, 49);\\
~~~~~~$\_$mm256$\_$store$\_$pd ($\&$A[j+0][k], r0);\\
~~~~~~// Store A[j+1][k], A[j+2][k], A[j+3][k]\\
~~~~$\rbrace$\\
~~$\rbrace$\\
$\rbrace$
\caption{Local assembly code generated by Firedrake for the Helmholtz problem when \emph{op-vect} is applied on top of the optimizations shown in Figure~\ref{code:helmholtz-licm}. Here, we assume the polynomial order is $p=2$, since \emph{op-vect} can not be used when an iteration space dimension is smaller than the vector length. The original size of the $j$-$k$ iteration space (i.e. before padding was applied) was 6$\times$6. In this example, the unroll-and-jam factor is 1.}
\label{code:helmholtz-opvect}
\end{algorithm}

@@

\begin{algorithm}[t]
\small
\KwSty{void} helmholtz(double A[3][4], double **coords) $\lbrace$\\
%~~\KwSty{$\#$define} ALIGN $\_\_$attribute$\_\_$((aligned(32))) \\
~~~~// Same code as in Figure~\ref{code:helmholtz-licm} up to line 15 \\
%~~// Declaration of basis function matrices \\
%~~\\
%~~\KwSty{for} (\KwSty{int} i = 0; i$<$3; i++) $\lbrace$ \\
%~~~~double LI$\_$0[4] ALIGN;\\
%~~~~double LI$\_$1[4] ALIGN;\\
%~~~~\KwSty{for} (\KwSty{int} r = 0; r$<$4; r++) $\lbrace$ \\
%~~~~~~LI$\_$0[r] = ((K1*X$\_$D10[i][r])+(K3*X$\_$D01[i][r]));\\
%~~~~~~LI$\_$1[r] = ((K0*X$\_$D10[i][r])+(K2*X$\_$D01[i][r]));\\
%~~~~$\rbrace$\\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$3; j++) \\
~~~~~~\KwSty{$\#$pragma vector aligned}\\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k++) \\
~~~~~~~~A[j][k] += (Y[i][k]*Y[i][j]+LI$\_$0[k]*LI$\_$0[j])*det*W3[i];\\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$3; j++) \\
~~~~~~\KwSty{$\#$pragma vector aligned}\\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k++) \\
~~~~~~~~A[j][k] += LI$\_$1[k]*LI$\_$1[j]*det*W3[i];\\
$\rbrace$
\caption{Local assembly code generated by Firedrake for the Helmholtz problem when \emph{split} is applied on top of the optimizations shown in Figure~\ref{code:helmholtz-licm}. In this example, the split factor is 2.}
\label{code:helmholtz-split}
\end{algorithm}

@@

\begin{algorithm}[t]
  \textbf{The COFFEE Compiler}\\
  \KwIn{ast, wrapper, isa}
  \KwOut{code}
// Analyze ast and build optimization plan \\
it\_space = analyze(ast) \\
\If{\KwSty{not} it\_space}{
  ast.apply\_inter\_kernel\_vectorization(wrapper, isa)
  \Return{wrapper + ast.from\_ast\_to\_c()}
}
plan = cost\_model(it\_space.n\_inner\_arrays, isa.n\_regs) \\
// Optimize ast based on plan \\
ast.licm() \\
ast.padding() \\
ast.data\_align() \\
\If{plan.permute}{
  ast.permute\_assembly\_loops()
}
\If{plan.sz\_split}{
  ast.split(plan.sz\_split)
}
\If{plan.uaj\_factor}{
  uaj = MIN(plan.uaj\_factor, $\lfloor$it\_space.j.size/isa.vf$\rfloor$) \\
  ast.op\_vect(uaj)
}
\Return{wrapper + ast.from\_ast\_to\_c()}
\caption{Pseudocode of the COFFEE pipeline.}
\label{algo:PyOP2Compiler}
\end{algorithm}

@@

\subsection{Impact of Generalized Loop-invariant Code Motion}
\newcommand{\massmatrixtwodresultsnorms}{
\begin{tabularx}{1.0\textwidth}{p{3.5cm}p{3.5cm}|p{2.25cm}p{2.25cm}p{2.25cm}p{2.25cm}|}
\cline{3-6}
& & \multicolumn{2}{c|}{\texttt{Sandy Bridge}} & \multicolumn{2}{c|}{\texttt{Xeon Phi}}\\
\cline{3-6}
& & \multicolumn{1}{c}{\texttt{licm}} & \multicolumn{1}{c|}{\texttt{licm-ap}} & \multicolumn{1}{c}{\texttt{licm}} & \multicolumn{1}{c|}{\texttt{licm-ap}} \\
\cline{1-6}
%\multicolumn{1}{|c|}{\texttt{Helmholtz}} & \multicolumn{1}{c|}{\texttt{triangle}} & ~~1.34$\times$-4.32$\times$ & ~1.34$\times$-4.32$\times$ & ~~1.34$\times$-4.32$\times$ & ~1.34$\times$-4.32$\times$ \\
\multicolumn{1}{|c|}{\texttt{Helmholtz}} & \multicolumn{1}{c|}{\texttt{triangle}} & ~~1.05$\times$-1.68$\times$ & \multicolumn{1}{c|}{~1.32$\times$-4.14$\times$} & ~~1.15$\times$-1.70$\times$ & ~~1.54$\times$-3.27$\times$ \\
\multicolumn{1}{|c|}{\texttt{Helmholtz}} & \multicolumn{1}{c|}{\texttt{tetrahedron}} & ~~1.36$\times$-2.64$\times$ & \multicolumn{1}{c|}{~1.35$\times$-3.32$\times$} & ~~1.18$\times$-1.87$\times$ & ~~1.27$\times$-2.92$\times$ \\
\multicolumn{1}{|c|}{\texttt{Helmholtz}} & \multicolumn{1}{c|}{\texttt{prism}} & ~~2.16$\times$-2.45$\times$ & \multicolumn{1}{c|}{~2.42$\times$-2.63$\times$} & ~~0.96$\times$-1.64$\times$ & ~~1.84$\times$-1.72$\times$ \\
\cline{1-6}
\multicolumn{1}{|c|}{\texttt{Diffusion}} & \multicolumn{1}{c|}{\texttt{triangle}} & ~~1.09$\times$-1.64$\times$ & \multicolumn{1}{c|}{~1.37$\times$-4.28$\times$} & ~~1.13$\times$-1.22$\times$ & ~~0.92$\times$-2.79$\times$ \\
\multicolumn{1}{|c|}{\texttt{Diffusion}} & \multicolumn{1}{c|}{\texttt{tetrahedron}} & ~~1.30$\times$-3.11$\times$ & \multicolumn{1}{c|}{~1.41$\times$-3.82$\times$} & ~~1.08$\times$-2.11$\times$ & ~~1.10$\times$-5.53$\times$ \\
\multicolumn{1}{|c|}{\texttt{Diffusion}} & \multicolumn{1}{c|}{\texttt{prism}} & ~~1.81$\times$-2.70$\times$ & \multicolumn{1}{c|}{~2.55$\times$-3.13$\times$} & ~~1.04$\times$-2.74$\times$ & ~~1.73$\times$-4.14$\times$ \\
\cline{1-6}
\multicolumn{1}{|c|}{\texttt{Burgers}} & \multicolumn{1}{c|}{\texttt{triangle}} & ~~1.53$\times$-2.46$\times$ & \multicolumn{1}{c|}{~1.56$\times$-2.77$\times$} & ~~1.19$\times$-2.97$\times$ & ~~2.70$\times$-4.27$\times$ \\
\multicolumn{1}{|c|}{\texttt{Burgers}} & \multicolumn{1}{c|}{\texttt{tetrahedron}} & ~~1.58$\times$-2.24$\times$ & \multicolumn{1}{c|}{~1.59$\times$-2.10$\times$} & ~~1.03$\times$-1.21$\times$ & ~~1.31$\times$-1.55$\times$ \\
\multicolumn{1}{|c|}{\texttt{Burgers}} & \multicolumn{1}{c|}{\texttt{prism}} & ~~1.32$\times$-2.11$\times$ & \multicolumn{1}{c|}{~1.42$\times$-2.31$\times$} & ~~1.04$\times$-1.40$\times$ & ~~1.27$\times$-2.18$\times$ \\
\cline{1-6}
\end{tabularx}
}

\begin{table*}[t]\small
\massmatrixtwodresultsnorms
~
\caption{Performance improvement due to generalized loop-invariant code motion (\emph{licm} column) on the Helmholtz, Diffusion and Burgers problems, varying the element shape (triangle, tetrahedron, prism) and the polynomial order ($p \in [1, 4]$), using the Lagrange element family, over the original, non-optimized code. Each entry indicates the minimum and maximum speedup obtained in the range of $p$ values. The column \emph{licm-ap} illustrates the combination of \emph{licm} with data alignment and padding. Results are shown for both the Sandy Bridge and the Xeon Phi architectures.}
\label{table:perf-results-licm}
\end{table*}

>>>>

\begin{algorithm}[t]
\footnotesize
\KwSty{void} helmholtz(double A[3][3], double **coords) $\lbrace$\\
~~// K, det = Compute Jacobian (coords) \\
~~\\
~~\KwSty{static const double} W3[3] = $\lbrace$...$\rbrace$\\
~~\KwSty{static const double} X$\_$D10[3][3] = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\KwSty{static const double} X$\_$D01[3][3] = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$3; i++) \\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$3; j++) \\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$3; k++) \\
~~~~~~~~A[j][k] += ((Y[i][k]*Y[i][j]+\\
~~~~~~~~~~~+((K1*X$\_$D10[i][k]+K3*X$\_$D01[i][k])*\\
~~~~~~~~~~~*(K1*X$\_$D10[i][j]+K3*X$\_$D01[i][j]))+\\
~~~~~~~~~~~+((K0*X$\_$D10[i][k]+K2*X$\_$D01[i][k])*\\
~~~~~~~~~~~*(K0*X$\_$D10[i][j]+K2*X$\_$D01[i][j])))*\\
~~~~~~~~~~~*det*W3[i]);\\
$\rbrace$
\caption{Local assembly code generated by Firedrake for a Helmholtz problem on a 2D triangular mesh using Lagrange $p=1$ elements.}
\label{code:helmholtz}
\end{algorithm}

\begin{algorithm}[t]
\footnotesize
\KwSty{void} burgers(double A[12][12], double **c, double **w) $\lbrace$\\
~~// K, det = Compute Jacobian (c) \\
~~\\
~~\KwSty{static const double} W5[5] = $\lbrace$...$\rbrace$\\
~~\KwSty{static const double} X1$\_$D001[5][12] = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\KwSty{static const double} X2$\_$D001[5][12] = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~//11 other basis functions definitions.\\
~~...\\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$5; i++) $\lbrace$\\
~~~~\KwSty{double} F0 = 0.0;\\
~~~~//10 other declarations (F1, F2,...)\\
~~~~...\\
~~~~\KwSty{for} (\KwSty{int} r = 0; r$<$12; r++) $\lbrace$\\
~~~~~~F0 += (w[r][0]*X1$\_$D100[i][r]);\\
~~~~~~//10 analogous statements (F1, F2, ...)\\
~~~~$\rbrace$\\
~~~~...\\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$12; j++) \\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$12; k++) \\
~~~~~~~~A[j][k] += (..(K5*F9)+(K8*F10))*Y1[i][j])+\\
~~~~~~~~~+(((K0*X1$\_$D100[i][k])+(K3*X1$\_$D010[i][k])+\\
~~~~~~~~~+(K6*X1$\_$D001[i][k]))*Y2[i][j]))*F11)+\\
~~~~~~~~~+(..((K2*X2$\_$D100[i][k])+...+(K8*X2$\_$D001[i][k])*\\
~~~~~~~~~*(K2*X2$\_$D100[i][j])+...+(K8*X2$\_$D001[i][j]))..)+\\
~~~~~~~~~+ $<$roughly a hundred sum/muls go here$>$)..)*\\
~~~~~~~~~*det*W5[i]);\\
~~$\rbrace$ \\
$\rbrace$
\caption{Local assembly code generated by Firedrake for a Burgers problem on a 3D tetrahedral mesh using Lagrange $p=1$ elements.}
\label{code:burgers}
\end{algorithm}

@@

\begin{algorithm}[t]
\footnotesize
\KwSty{void} helmholtz(double A[3][4], double **coords) $\lbrace$\\
~~\KwSty{$\#$define} ALIGN $\_\_$attribute$\_\_$((aligned(32))) \\
~~// K, det = Compute Jacobian (coords) \\
~~\\
~~\KwSty{static const double} W3[3] ALIGN = $\lbrace$...$\rbrace$\\
~~\KwSty{static const double} X$\_$D10[3][4] ALIGN = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\KwSty{static const double} X$\_$D01[3][4] ALIGN = $\lbrace\lbrace$...$\rbrace\rbrace$\\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$3; i++) $\lbrace$ \\
~~~~double LI$\_$0[4] ALIGN;\\
~~~~double LI$\_$1[4] ALIGN;\\
~~~~\KwSty{for} (\KwSty{int} r = 0; r$<$4; r++) $\lbrace$ \\
~~~~~~LI$\_$0[r] = ((K1*X$\_$D10[i][r])+(K3*X$\_$D01[i][r]));\\
~~~~~~LI$\_$1[r] = ((K0*X$\_$D10[i][r])+(K2*X$\_$D01[i][r]));\\
~~~~$\rbrace$\\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$3; j++) \\
~~~~~~\KwSty{$\#$pragma vector aligned}\\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k++) \\
~~~~~~~~A[j][k] += (Y[i][k]*Y[i][j]+LI$\_$0[k]*LI$\_$0[j]+\\
~~~~~~~~~~~~~~~~~~~~+LI$\_$1[k]*LI$\_$1[j])*det*W3[i]);\\
~~$\rbrace$\\
$\rbrace$
\caption{Local assembly code generated by Firedrake when padding, data alignment, and \emph{licm} are applied to the Helmholtz problem given in Figure~\ref{code:helmholtz} for an AVX architecture. In this specific case, sub-expressions invariant to $j$ are identical to those invariant to $k$, so they can be precomputed once in the $r$ loop.}
\label{code:helmholtz-licm}
\end{algorithm}

@@

\begin{algorithm}[t]
\footnotesize
\KwSty{void} helmholtz(double A[8][8], double **coords) $\lbrace$\\
~~// K, det = Compute Jacobian (coords) \\
~~// Declaration of basis function matrices \\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$6; i++) $\lbrace$ \\
~~~~// Do loop-invariant code motion \\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$4; j+=4) \\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$8; k+=4) $\lbrace$\\
~~~~~~~~// $load$ and $set$ intrinsics \\
~~~~~~~~// Compute A[1,1],A[2,2],A[3,3],A[4,4] \\
~~~~~~~~// One $permute\_pd$ intrinsics per $k$-loop $load$\\
~~~~~~~~// Compute A[1,2],A[2,1],A[3,4],A[4,3] \\
~~~~~~~~// One $permute2f128\_pd$ intrinsics per $k$-loop $load$\\
~~~~~~~~// ...\\
~~~~~~$\rbrace$\\
~~~~// Remainder loop (from $j=4$ to $j=6$)\\
~~$\rbrace$\\
~~// Restore the storage layout:\\
~~\KwSty{for} (\KwSty{int} j = 0; j$<$4; j+=4) $\lbrace$\\
~~~~$\_\_$m256d r0, r1, r2, r3, r4, r5, r6, r7;\\
~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$8; k+=4) $\lbrace$\\
~~~~~~r0 = $\_$mm256$\_$load$\_$pd ($\&$A[j+0][k]);\\
~~~~~~// Load A[j+1][k], A[j+2][k], A[j+3][k]\\
~~~~~~r4 = $\_$mm256$\_$unpackhi$\_$pd (r1, r0);\\
~~~~~~r5 = $\_$mm256$\_$unpacklo$\_$pd (r0, r1);\\
~~~~~~r6 = $\_$mm256$\_$unpackhi$\_$pd (r2, r3);\\
~~~~~~r7 = $\_$mm256$\_$unpacklo$\_$pd (r3, r2);\\
~~~~~~r0 = $\_$mm256$\_$permute2f128$\_$pd (r5, r7, 32);\\
~~~~~~r1 = $\_$mm256$\_$permute2f128$\_$pd (r4, r6, 32);\\
~~~~~~r2 = $\_$mm256$\_$permute2f128$\_$pd (r7, r5, 49);\\
~~~~~~r3 = $\_$mm256$\_$permute2f128$\_$pd (r6, r4, 49);\\
~~~~~~$\_$mm256$\_$store$\_$pd ($\&$A[j+0][k], r0);\\
~~~~~~// Store A[j+1][k], A[j+2][k], A[j+3][k]\\
~~~~$\rbrace$\\
~~$\rbrace$\\
$\rbrace$
\caption{Local assembly code generated by Firedrake for the Helmholtz problem when \emph{op-vect} is applied on top of the optimizations shown in Figure~\ref{code:helmholtz-licm}. Here, we assume the polynomial order is $p=2$, since \emph{op-vect} can not be used when an iteration space dimension is smaller than the vector length. The original size of the $j$-$k$ iteration space (i.e. before padding was applied) was 6$\times$6. In this example, the unroll-and-jam factor is 1.}
\label{code:helmholtz-opvect}
\end{algorithm}

@@

\begin{algorithm}[t]
\footnotesize
\KwSty{void} helmholtz(double A[3][4], double **coords) $\lbrace$\\
~~~~// Same code as in Figure~\ref{code:helmholtz-licm} up to line 15 \\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$3; j++) \\
~~~~~~\KwSty{$\#$pragma vector aligned}\\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k++) \\
~~~~~~~~A[j][k] += (Y[i][k]*Y[i][j]+LI$\_$0[k]*LI$\_$0[j])*det*W3[i];\\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$3; j++) \\
~~~~~~\KwSty{$\#$pragma vector aligned}\\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k++) \\
~~~~~~~~A[j][k] += LI$\_$1[k]*LI$\_$1[j]*det*W3[i];\\
$\rbrace$
\caption{Local assembly code generated by Firedrake for the Helmholtz problem when \emph{split} is applied on top of the optimizations shown in Figure~\ref{code:helmholtz-licm}. In this example, the split factor is 2.}
\label{code:helmholtz-split}
\end{algorithm}

@@

\begin{algorithm}[t]
\footnotesize
  \textbf{The COFFEE Compiler}\\
  \KwIn{ast, wrapper, isa}
  \KwOut{code}
// Analyze ast and build optimization plan \\
it\_space = analyze(ast) \\
\If{\KwSty{not} it\_space}{
  ast.apply\_inter\_kernel\_vectorization(wrapper, isa)\\
  \Return{wrapper + ast.from\_ast\_to\_c()}
}
plan = cost\_model(it\_space.n\_inner\_arrays, isa.n\_regs) \\
// Optimize ast based on plan \\
ast.licm() \\
ast.padding() \\
ast.data\_align() \\
\If{plan.permute}{
  ast.permute\_assembly\_loops()
}
\If{plan.sz\_split}{
  ast.split(plan.sz\_split)
}
\If{plan.uaj\_factor}{
  uaj = MIN(plan.uaj\_factor, $\lfloor$it\_space.j.size/isa.vf$\rfloor$) \\
  ast.op\_vect(uaj)
}
\Return{wrapper + ast.from\_ast\_to\_c()}
\caption{Pseudocode of the COFFEE pipeline.}
\label{algo:PyOP2Compiler}
\end{algorithm}

@@

\subsection{Impact of Generalized Loop-invariant Code Motion}
%\newcommand{\massmatrixtwodresultsnorms}{
%\begin{tabularx}{1.0\textwidth}{p{3.5cm}p{3.5cm}|p{2.25cm}p{2.25cm}p{2.25cm}p{2.25cm}|}
%\cline{3-6}
%& & \multicolumn{2}{c|}{\texttt{Sandy Bridge}} & \multicolumn{2}{c|}{\texttt{Xeon Phi}}\\
%\cline{3-6}
%& & \multicolumn{1}{c}{\texttt{licm}} & \multicolumn{1}{c|}{\texttt{licm-ap}} & \multicolumn{1}{c}{\texttt{licm}} & \multicolumn{1}{c|}{\texttt{licm-ap}} \\
%\cline{1-6}
%%\multicolumn{1}{|c|}{\texttt{Helmholtz}} & \multicolumn{1}{c|}{\texttt{triangle}} & ~~1.34$\times$-4.32$\times$ & ~1.34$\times$-4.32$\times$ & ~~1.34$\times$-4.32$\times$ & ~1.34$\times$-4.32$\times$ \\
%\multicolumn{1}{|c|}{\texttt{Helmholtz}} & \multicolumn{1}{c|}{\texttt{triangle}} & ~~1.05$\times$-1.68$\times$ & \multicolumn{1}{c|}{~1.32$\times$-4.14$\times$} & ~~1.15$\times$-1.70$\times$ & ~~1.54$\times$-3.27$\times$ \\
%\multicolumn{1}{|c|}{\texttt{Helmholtz}} & \multicolumn{1}{c|}{\texttt{tetrahedron}} & ~~1.36$\times$-2.64$\times$ & \multicolumn{1}{c|}{~1.35$\times$-3.32$\times$} & ~~1.18$\times$-1.87$\times$ & ~~1.27$\times$-2.92$\times$ \\
%\multicolumn{1}{|c|}{\texttt{Helmholtz}} & \multicolumn{1}{c|}{\texttt{prism}} & ~~2.16$\times$-2.45$\times$ & \multicolumn{1}{c|}{~2.42$\times$-2.63$\times$} & ~~0.96$\times$-1.64$\times$ & ~~1.84$\times$-1.72$\times$ \\
%\cline{1-6}
%\multicolumn{1}{|c|}{\texttt{Diffusion}} & \multicolumn{1}{c|}{\texttt{triangle}} & ~~1.09$\times$-1.64$\times$ & \multicolumn{1}{c|}{~1.37$\times$-4.28$\times$} & ~~1.13$\times$-1.22$\times$ & ~~0.92$\times$-2.79$\times$ \\
%\multicolumn{1}{|c|}{\texttt{Diffusion}} & \multicolumn{1}{c|}{\texttt{tetrahedron}} & ~~1.30$\times$-3.11$\times$ & \multicolumn{1}{c|}{~1.41$\times$-3.82$\times$} & ~~1.08$\times$-2.11$\times$ & ~~1.10$\times$-5.53$\times$ \\
%\multicolumn{1}{|c|}{\texttt{Diffusion}} & \multicolumn{1}{c|}{\texttt{prism}} & ~~1.81$\times$-2.70$\times$ & \multicolumn{1}{c|}{~2.55$\times$-3.13$\times$} & ~~1.04$\times$-2.74$\times$ & ~~1.73$\times$-4.14$\times$ \\
%\cline{1-6}
%\multicolumn{1}{|c|}{\texttt{Burgers}} & \multicolumn{1}{c|}{\texttt{triangle}} & ~~1.53$\times$-2.46$\times$ & \multicolumn{1}{c|}{~1.56$\times$-2.77$\times$} & ~~1.19$\times$-2.97$\times$ & ~~2.70$\times$-4.27$\times$ \\
%\multicolumn{1}{|c|}{\texttt{Burgers}} & \multicolumn{1}{c|}{\texttt{tetrahedron}} & ~~1.58$\times$-2.24$\times$ & \multicolumn{1}{c|}{~1.59$\times$-2.10$\times$} & ~~1.03$\times$-1.21$\times$ & ~~1.31$\times$-1.55$\times$ \\
%\multicolumn{1}{|c|}{\texttt{Burgers}} & \multicolumn{1}{c|}{\texttt{prism}} & ~~1.32$\times$-2.11$\times$ & \multicolumn{1}{c|}{~1.42$\times$-2.31$\times$} & ~~1.04$\times$-1.40$\times$ & ~~1.27$\times$-2.18$\times$ \\
%\cline{1-6}
%\end{tabularx}
%}
